---
title: "Survey Data Analysis with Kobocruncher"
subtitle: "Session 4 - Setting Crosstabulation"
author: "[Link to Documentation](https://edouard-legoupil.github.io/kobocruncher/) -- [Link to Previous Session](03-Grouping_Questions.html) -- [Link to Next Session](05-Searching_Asssociation.html)  "
date: "Training Content as of `r format(Sys.Date(),  '%d %B %Y')`"
output:
  unhcrdown::html_slides:
    self_contained: true
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  fig.showtext = TRUE
)
library(tidyverse)
library(unhcrthemes)
library(fontawesome)
```

 
## Cross-tabulation: what is that?


When you want to compare the results for one or more variables with the results of another, then you need to cross-tabulate


This will allow to examine relationships within the data that might not be readily apparent when only looking at total survey responses.


---

## Frequent variable used for cross-tabulation

Typically cross-tabulation are required for standard disaggregation (i.e. breakdown) variables

 * Gender of the main respondent

 * Country of Origin

 * Area of residence in the country (here you might group together different areas based on frequency to obtain meaningfull representation -see session 7)
 
 * Age
 
 * Displacement type

---

## How can I crosstabulate questions within my survey?


By default the initial exploration report does not include cross-tabulation as it depends on your data:


 1. Open the expanded xlsform 
 
 2. set up yes in the `disaggregation` column for the variable you want to use for cross-tabulation
 
---

## What will you get then?

Once you regenerate the chart with this new setting in your analysis plan, every single questions will be be presented in additional charts for each `disaggregation` variable

 .bg-blue[
This setting will automatically inflate considerably the size of your exploration report.

You may save you expanded `xlsform` with a new name and adjust your parameters settings so that you can create multiple exploration report exploring different `disaggregation` options

Do not forget, at this stage you are exploring your data so that you can select the final charts you want to include in the presentation for the joint data interpretation session

]   

---

## What happenned?


The function [kobo_cruncher](https://edouard-legoupil.github.io/kobocruncher/reference/kobo_cruncher.html) then loop around all the questions, and in adition of the default simple tabulations, and based on the data types, generate standard charts:

 * `select_one` simple bar chart with [plot_select_one_cross](https://edouard-legoupil.github.io/kobocruncher/reference/plot_select_one_cross.html)
 
 * `select_multiple` bar chart [plot_select_multiple_cross](https://edouard-legoupil.github.io/kobocruncher/reference/plot_select_multiple_cross.html)
 
 * `integer` or `numeric` or `date` histogram with [plot_integer_cross](https://edouard-legoupil.github.io/kobocruncher/reference/plot_integer_cross.html)
  
 

---

class: inverse, center, middle

# TIME TO PRACTISE ON YOUR OWN! 

### .large[.white[`r fontawesome::fa("fas fa-pen-nib"  )  `] **5 minutes! **] 

 
```{r echo=FALSE}
countdown::countdown(minutes = 5,
                     seconds = 0,
                     margin = "5%",
                     font_size = "8em",
                     style = "position: relative; width: min-content;",
                     # Fanfare when it's over
                     play_sound = TRUE)
```

Open your xlsform and fill in the .large[disaggregation] column, save and knit again your report
 


Do not hesitate to raise your questions in the [ticket system](https://github.com/Edouard-Legoupil/kobocruncher/issues/new) or in the chat so the training content can be improved accordingly! 
 
---
class: inverse, center, middle



### .large[.white[`r fontawesome::fa("fas fa-clock"  )  `] **Let's take a break! **] 
 
```{r echo=FALSE}
countdown::countdown(minutes = 5,
                     seconds = 0,
                     margin = "5%",
                     font_size = "8em",
                     style = "position: relative; width: min-content;",
                     # Fanfare when it's over
                     play_sound = TRUE)
```


__Next session__:   

[05-Searching_Asssociation: beyond the cross-tabulation, it is key to check how variables relate to each other. This can be done with statistical test that verify if variables are significantly associated](05-Searching_Asssociation.html) 
