---
title: "Survey Data Analysis with Kobocruncher"
subtitle: "Session 5 - Searching for Association"
author: "[Link to Documentation](https://edouard-legoupil.github.io/kobocruncher/) -- [Link to Previous Session](04-Setting_Crosstabulation.html) -- [Link to Next Session](06-Cleaning_and_Indicator_calculation.html)  "
date: "Training Content as of `r format(Sys.Date(),  '%d %B %Y')`"
output:
  unhcrdown::html_slides:
    self_contained: true
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  fig.showtext = TRUE
)
library(tidyverse)
library(unhcrthemes)
library(fontawesome)
```


## Significant associations & Chi-square 

The Chi-square statistic is the primary statistic used for testing the statistical significance of the cross-tabulation table. 

Chi-square tests determine whether or not the two variables are independent:

 * If the variables are independent (have no relationship), then the results of the statistical test will be “non-significant” and we are not able to reject the null hypothesis, meaning that we believe there is no relationship between the variables. 
 
 * If the variables are related, then the results of the statistical test will be “statistically significant” and we can reject the null hypothesis, meaning that we can state that there is some relationship between the variables.


---

## How to interpret correlation tables? 



Correlation tables only appear if the question has a significant relationship with another indicator. All the answer options for the two questions are seen in the rows and columns and each cell conveys the relationship between the two answer options. 

Each cell value indicates two characteristics of the relationship: the strength (numeric value) and the direction (- or + sign). 

The larger the numeric value, the stronger the relationship. 


---

## Let's look at an example



In this example the strongest relationships are identified by the bigger numbers and circles. 

Choose the strongest or most relevant positive (+) relationship to ease the interpretation. 

For this table, the following interpretation can be made: __“When a person reports that they have witnessed or suffered any incidents (Row ‘Yes’), they are more likely to report that they feel lonely or socially isolated in their current location (Column ‘Strongly Agree’)__.



---
class: inverse, center, middle

# TIME TO PRACTISE ON YOUR OWN! 

### .large[.white[`r fontawesome::fa("fas fa-pen-nib"  )  `] **5 minutes! **] 

 
```{r echo=FALSE}
countdown::countdown(minutes = 5,
                     seconds = 0,
                     margin = "5%",
                     font_size = "8em",
                     style = "position: relative; width: min-content;",
                     # Fanfare when it's over
                     play_sound = TRUE)
```


 - Download again locally and fill in the correlation
 
 - upload and knit again your report
 

Do not hesitate to raise your questions in the [ticket system](https://github.com/Edouard-Legoupil/kobocruncher/issues/new) or in the chat so the training content can be improved accordingly! 
 
---
class: inverse, center, middle



### .large[.white[`r fontawesome::fa("fas fa-clock"  )  `] **Let's take a break! **] 
 
```{r echo=FALSE}
countdown::countdown(minutes = 5,
                     seconds = 0,
                     margin = "5%",
                     font_size = "8em",
                     style = "position: relative; width: min-content;",
                     # Fanfare when it's over
                     play_sound = TRUE)
```


__Next session__: 

[06-Cleaning_and_Indicator_calculation after reviewing the first iteration of the reports, we can identify potential data quality issue to fix and create calculated variable and indicator ](06-Cleaning_and_Indicator_calculation.html)
