---
title: "Survey Data Analysis with Kobocruncher"
subtitle: "Session 8 - Weighting"
author: "[Link to Documentation](https://edouard-legoupil.github.io/kobocruncher/) -- [Link to Previous Session](07-Anonymising.html) -- [Link to Next Session](09-Publishing.html)  "
date: "Training Content as of `r format(Sys.Date(),  '%d %B %Y')`"
output:
  unhcrdown::html_slides:
    self_contained: true
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  fig.showtext = TRUE
)
library(tidyverse)
library(unhcrthemes)
library(fontawesome)
```


## Review of sampling designs

These features can be combined to form one design

 *  __Simple random sampling__: every unit has the same chance of being selected
      *   Without replacement: units can only be selected once
      *   With replacement: units can be selected more than once

 *  __Systematic sampling__: sample n individuals from a ordered list and sampling individuals at an interval with a random starting point

 *  __Probability proportional to size__: probability of selection is proportional to "size"

 *  __Stratified sampling__: divide population into mutually exclusive subgroups (strata). Randomly sample within each stratum

 *  __Clustered sampling__: divide population into mutually exclusive subgroups (clusters). Randomly sample clusters and then individuals within clusters

---

## Determining the design

Look at documentation associated with the analysis file

Keywords to look for: methodology, design, analysis guide, technical documentation

Documentation will indicate the variables needed to specify the design. Look for:

 *  weight (almost always)
 *  strata and/or clusters/PSUs. Sometimes pseudo-strata and pseudo-cluster OR
 *  replicate weights (this is used instead of strata/clusters for analysis)
 *  might also see finite population correction or population sizes

---

##  Syntax for common designs



```{r eval=FALSE, include=TRUE}
# simple random sample (SRS)
apisrs %>% as_survey_design(fpc = fpc)
# stratified sample
apistrat %>% as_survey_design(strata = stype, weights = pw)
# one-stage cluster sample
apiclus1 %>% as_survey_design(ids = dnum, weights = pw, fpc = fpc)
# two-stage cluster sample, weights computed from pop size
apiclus2 %>% as_survey_design(ids = c(dnum, snum), fpc = c(fpc1, fpc2))
# stratified, cluster design
apistrat %>% as_survey_design(ids = dnum, strata = stype, weights =pw, nest = TRUE)
```

---
class: inverse, center, middle

# Thank you


__Next session__:   [09-Publishing](9-Publishing.html) The final stage of the process implies to facilitate both the interpretation of the data and the creation of data story telling content 
